---
title: "report"
author: "huangjiushi"
date: "2016年3月28日"
output: html_document
---
IS4250 Health Care Analysis Report


An Exploratory Analysis of Liver Patients from USA and INDIA





Group 08:	Huang Jiushi (A0105542R)
			    Ye Shujian  (A0


1.     Introduction
1.1  Background
According to the World Health Organization (WHO) report published in 2014, the number of Liver Disease Deaths in India has reached 216,865 or 2.44% of total deaths. The age adjusted Death Rate is 21.96 per 100,000 of population which made Liver Disease become one of the top 10 causes of death in India. In order to improve the liver medical diagnosis accuracy rate, Prof Tekkali, from Aditya Institute of Technology and Management in India, performed several researches based on different data sets. Significant differences in classification accuracy of various classifiers with different data sets have been indicated. In order to envisage the reason for the difference between different data sets to further improve the diagnosis accuracy, an exploratory analysis was carried out.
 
1.2  Project Objectives
In order to improve the diagnosis efficiency, India is currently developing the automatic medical diagnosis to instead the manual diagnosis methods. With the purpose of improving the system accuracy, the results of this study are expected to help the development of the automatic medical diagnosis system and the needs for its localization settings based on the geographical region.
 
1.3  Data Set Introduction
The study contains 2 data sets, taken from the University of California at Irvine (UCI) Machine Learning Repository and a local hospital in north east, Andhra Pradesh, India respectively. The reason why they choose one data set from USA and one from India is because significant differences within these two data sets are expected to be found due to the regional disparity.
 
The first data set indicates UCI data set which contains 345 records with 6 attributes.
145 of them are labelled as liver patients and others are not. The attributes of the UCI data set are MCV(mean corpuscular volume) , Alkphos(alkaline phosphotase), SGPT(alamine aminotransferase), SGOT(aspartate aminotransferase), Gammagt(gamma-glutamyl transpeptidase) and Drinks(number of half-pint equivalents of alcoholic beverages drunk per day).
 
The second data set indicates India data set which has 583 liver patient records with 10 attributes including Gender, Age, Total_Bilirubin, Direct_Bilirubin, Total_Protiens, Albumin, A/G_Ratio(Albumin and Globulin Ratio), SGPT(alamine aminotransferase), SGOT(aspartate aminotransferase) and Alkphos(alkaline phosphotase). Similarly, 416 records of India data set are labelled as liver patients and others are considered as non-liver patients. The common attributes from the two data sets are Alkphos, SGPT and SGOT, which will be taken for the purpose of comparison.
 
2.     Methodology
 
Standard statistical methods of One-way Analysis of Variance (ANOVA) and Multivariate Analysis of Variance (MANOVA) are applied to evaluate the significance between two data sets for better classification.
 
ANOVA is a collection of statistical models used to analyze the differences in a single dependent variable among group means and their associated procedure. It will compare the means between the groups and determine whether any of those means are significantly different from each other. Specifically, it tests the null hypothesis:
H0:  µ1 = µ2 = µ3 = … = µk
Where µ = group and k = number of groups. If the one-way ANOVA returns a significant result, the alternative hypothesis will be accepted, which means that
there are at least 2 group means that are significantly different from each other.
 
MANOVA is a technic to compare multivariate sample means. As a multivariate procedure, it is used when there are two or more dependent variables, and is typically followed by significance tests involving individual dependent variables separately. It can help find the relationship among the dependent variables and independent variables respectively. The purpose of the MANOVA is to look for an effect of one or more independent variables on several dependent variables at the same time.

 

3. Related Work



4. Result and Discussion

There are three common attributes between the Indian data set and the US data set. They are SGOT, SGPT, Alkphos and these three attributes are integers. To show that differences between Indian liver patients and US liver patients exist, these three common attributes will be the only measures.

The three experiments will be conducted on three different samples. For each experiment, we will have 7 comparisons to show that the significant differences exist. First three comparisons will examine whether significant differences on individual common attribute exist between India patients and US patients, i.e. comparison on SGOT, comparison on SGPT, and comparison on Alkphos. The next three comparisons will examine whether significant differences exist on the combination of two common attributes between India patients and US patients, i.e. comparison on SGOT and SGPT, comparison on SGOT and Alkphos, comparison on SGPT and Alkphos. The last comparison will examine whether significant differences on the combination of all three common variables exist between Indian patients and US patients, i.e. comparison on SGOT, SGPT, and Alkphos. The first three comparisons will use ANOVA because we only need to examine the difference of data sets on single variable. The rest of the comparisons will use MANOVA as we need to examine the difference of data sets on multiple variables.

4.1 Experiment 1
For the first experiment, our sample will be all the records in the two data sets which means both liver patients and non-liver patients are included in the experiment. 

The first comparison will be on Alkphos using ANOVA. We calculated that the mean of Alkphos of US patients is 69.87 and the mean of Alkphos of Indian patients is 290.58. Then we can calculate the sum of squares of Alkphos between two groups (Indian patients and US patients) which is 10557739.946. The numerator degree of freedom is 1 and it is equal to number of groups minus 1. Hence, we can get the mean square between two groups which is 10557739.946 and it is equal to sum of squares between two groups divided by the numerator degree of freedom.

For the sum of squares within two groups, it is equal to 34464783.484. The denominator degree of freedom is 926 and it is equal to the total number of records in two groups (928) minus number of groups. Hence we can get the mean square within two groups which is 37218.989 and it is equal to sum of squares within two groups divided by the denominator degree of freedom.

After knowing the value of the mean square between two groups and the mean square within two groups we can calculate the F-ratio which is equal to mean square between two groups divided by mean square within two groups, and it is equal to 283.665. We can use the numerator degree of freedom and the denominator degree of freedom to look for the corresponding critical value in the F distribution table. The F distribution table we use is for 0.001 significance level. We found that the critical value for a numerator degree of freedom of 1 and a denominator degree of freedom of 926 and the 0.001 significance level is 10.82756617. Obviously, the F-ratio is much larger than the critical value. Hence, we safely conclude that the significant difference on the Alkphos attribute between Indian patients group and US patients group exist.

The second comparison will be on SGPT using ANOVA. The means of US patients group and Indian patients group are 30.41 and 80.71 respectively. The sum of squares between two groups and sum of squares within two groups are 548541.541 and 19540784.351 respectively. The numerator degree of freedom and denominator degree of freedom are the same as those in the first comparison and hence the critical value for 0.001 significance level is the same the critical value in the first comparison which is 10.82756617. The calculated F-ratio of this comparison is 25.994 which is greater than the critical value for 0.001 significance level. Hence, we safely conclude that the significant difference on the SGPT attribute between Indian patients group and US patients group exist.

The third comparison will be on SGOT using ANOVA. The means of US patients group and Indian patients group are 24.64 and 109.91 respectively. The sum of squares between two groups and sum of squares within two groups are 1575814.094 and 48616664.510 respectively. The numerator degree of freedom and denominator degree of freedom are the same as those in the first comparison and hence the critical value for 0.001 significance level is the same the critical value in the first comparison which is 10.82756617. The calculated F-ratio of this comparison is 30.014 which is greater than the critical value for 0.001 significance level. Hence, we safely conclude that the significant difference on the SGOT attribute between Indian patients group and US patients group exist.

5. Conclusion

6. References
```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```


